### 2023-05-13 talk/slides.Rmd

---
title: "Curation of NYC 311 Request Open Data"
author: ""     # to be supplied through headers-includes below
institute: ""  # to be supplied through headers-includes below
output:
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    slide_level: 2
#    toc: true
    number_sections: false
classoption: "aspectratio=169"
header-includes:
  - \usepackage{amsmath, amsfonts, bm}
  - \usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{multirow}
  - \usepackage{adjustbox}
  - \setbeamercolor{frametitle}{bg=white}
  - \AtBeginSection[]
    {
    \begin{frame}<beamer>
      \frametitle{Outline}
      \tableofcontents[currentsection]
    \end{frame}
    }
  - \AtBeginDocument
    {
    \title[NYC 311 Requests]{Curation of NYC 311 Request Open Data}
    \author[Jun Yan]{David Tussey\inst{1} and Jun Yan\inst{2}}
    \institute[University of Connecticut]{\inst{1} Independent Researcher\\
    \inst{2} University of Connecticut}
    }
bibliography: ["../manuscript/ref.bib"]
link-citations: true
link-color: true
# csl: apa.csl
---

# Introduction

## NYC 311 Open Data

- Open data is great ... trend ...
- NYC is a forerunner in open data ...
- 311 request data is one of the most popular open data at NYC Open Data
- Found some issues from teaching ...

# Issues with NYC 311 Requests Data

## Data Structure

Demonstration with data for the week of Januare 15, 2023, 
downloaded on February 22, 2023.
```{r read}
d311 <- read.csv("../data/nyc311_011523-012123_by022023.csv", header = TRUE)
names(d311) <- tolower(gsub("\\.", "_", names(d311)))

names(d311)
```

## Invalid Created/Closed Date
```{r date}
t1 <- strptime(d311$closed_date, format = '%m/%d/%Y %I:%M:%S %p')
t0 <- strptime(d311$created_date, format = '%m/%d/%Y %I:%M:%S %p')
tt <- as.numeric(difftime(t1, t0, units =  "secs"))
sum(tt < 0, na.rm = TRUE)
head(subset(d311, tt < 0, select = c("created_date", "closed_date")), 5)
```

## Invalid Zipcode

## Inconsistent Data

# Suggestions on Open Data Releasing Protocols

# References {.allowframebreaks} 
\scriptsize