### 2023-06-09 talk/slides.Rmd
went through slides with David

---
title: "Curation of NYC 311 Request Open Data"
author: ""     # to be supplied through headers-includes below
institute: ""  # to be supplied through headers-includes below
output:
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    slide_level: 2
#    toc: true
    number_sections: false
classoption: "aspectratio=169"
header-includes:
  - \usepackage{amsmath, amsfonts, bm}
  - \usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{multirow}
  - \usepackage{adjustbox}
  - \setbeamercolor{frametitle}{bg=white}
  - \AtBeginSection[]
    {
    \begin{frame}<beamer>
      \frametitle{Outline}
      \tableofcontents[currentsection]
    \end{frame}
    }
  - \AtBeginDocument
    {
    \title[NYC 311 Requests]{Curation of NYC 311 Request Open Data}
    \author[Jun Yan]{David Tussey\inst{1} and Jun Yan\inst{2}}
    \institute[University of Connecticut]{\inst{1} Former Executive Director, NYC DoITT\\
    \inst{2} University of Connecticut}
    }
bibliography: ["../manuscript/ref.bib"]
link-citations: true
link-color: true
# csl: apa.csl
---

```{r setup, include = FALSE, echo = FALSE}
library(stringr)
source("../code/functions.R")

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  paste0("\n \\", "small","\n\n", x, "\n\n \\normalsize")
})
## Load the 311 SR data file
```

# Introduction

## NYC 311 Open Data

- Open Data is viewed as a right for all.
- Cities lead open data efforts wtih >50 cities with formal programs.
- NYC is a forerunner with Open Data Law in 2012.
- Today NYC Open Data has 3511 datasets from 98 Agencies, accesible online.
- 311 Service Request data is the most viewed dataset at NYC Open Data.
- Used in STAT 3255/5255 Introduction to Data Science at UConn
- Exported 1-31 March 2023 311 SRs for analysis (~250K records).

# Issues with NYC 311 Requests Data

## Data Columns
<!-- - Column names: Remove trailing "."s.  Remove extra "."s -->
<!-- - Convert "." separator to "_".  Covert column names to lower cases -->
- Tidied column names (lower cases; underscore delimited; no trailing spaces)
- Last five columns not referenced in the Data Dictionary (not the case last year)
    + The NYC-OD Team advised these were "computed" fields derived from other data.

```{r read}
d311 <- read.csv("../data/311_202303by202305.csv", header = TRUE)
names(d311) <- makeColNamesUserFriendly(names(d311))
dim(d311)
tail(names(d311))
```

## Contributions by Agencies

```{r byAgency}
d311freq_agency <- table(d311$agency)
sort(d311freq_agency, decreasing = TRUE)
```

## Missing data: Character columns
- Various forms: "", "unspecified", and "unknown"
- Results show 3 groupings: all blanks (>90%), a few blanks (12-44%), or nearly no blanks (1-5%)

```{r blanks}
cMissingDataPerColumn <- charColumnsMissingData(d311)
cMissingDataPerColumn[order(-rowSums(cMissingDataPerColumn)), ]
```

## Missing data: Numeric Colummns

```{r missing}
nMissingDataPerColumn <- numColumnsMissingData(d311)
sort(nMissingDataPerColumn, decreasing = TRUE)
```

<!-- ## Unspecified/Unknown data columns -->
<!-- - Several columns have many "Unspecified" or "Unknown" values. These are deemed valid values. -->
<!-- ```{r uknown} -->
<!-- tail(missingDataPerColumn[ order( -missingDataPerColumn[ , 2] ), ], 4) -->
<!-- ``` -->

## Invalid Data: Created/Closed Date
```{r date}
t1 <- strptime(d311$closed_date, format = '%m/%d/%Y %I:%M:%S %p')
t0 <- strptime(d311$created_date, format = '%m/%d/%Y %I:%M:%S %p')
tt <- as.numeric(difftime(t1, t0, units =  "day"))
badTime <- subset(d311, tt <= 0,
                  select = c("created_date", "closed_date", "agency"))

## exactly the same created and closed date down to the second; a lot!
sum(tt == 0, na.rm = TRUE)
table(subset(badTime, created_date == closed_date, select = "agency"))
```

## Bad Closed Date
```{r baddate}
## quite some negative response times
table(subset(badTime, created_date != closed_date, select = "agency"))

summary(tt[tt < 0 & (!is.na(tt))])
```

## Invalid Data: Incident Zipcodes
```{r zip1}
validZipcodes <- scan("../data/NYCzipcodes.csv")

badIncident_zip <- subset(d311, ! incident_zip %in% validZipcodes,
                          select = c("incident_zip", "agency"))
head(sort(table(badIncident_zip$incident_zip), decreasing = TRUE))
```

## Invalid Data: Zipcodes (One of the last 5 columns)
```{r zip2}
badZipcode  <- subset(d311, ! zip_codes %in% validZipcodes,
                      select = c("zip_codes", "agency"))
nrow(badZipcode)  # mostly invalid!
```

## Invalid Data: Borough
```{r borough}
validBorough <- c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "STATEN ISLAND",
                  "Unspecified")
badBorough <- subset(d311, ! borough %in% validBorough,
                     select = c("borough", "agency"))
head(table(badBorough$agency))
head(table(subset(d311, borough == "Unspecified")$agency))
```

## Invalid Data: Borough_boundaries and Park_borough
```{r borough2}
validBB <- c("1", "2", "3", "4", "5")
badBB <- subset(d311, ! borough_boundaries %in% validBB,
                select = c("borough_boundaries", "agency"))
head(table(badBB$agency))

badPark_borough <- subset(d311, ! park_borough %in% validBorough,
                          select = c("park_borough", "agency"))
head(table(badPark_borough$agency))
```

## Invalid Data: Open Data Channel
```{r channel}
validChannel <- c("MOBILE", "ONLINE", "PHONE", "OTHER")
badChannel_type <- subset(d311, ! open_data_channel_type %in% validChannel,
                          select = c("open_data_channel_type", "agency"))
head(table(badChannel_type$agency)) ## UNKNOWN
```

## Invalid Data: Police Precincts
```{r nypd}
validPolice_precincts <- scan("../data/NYPDPrecincts2023.csv", skip = 1)
badPolice_precincts <- subset(
    d311, ! police_precincts %in% validPolice_precincts,
    select = c("police_precincts", "agency"))
sort(table(badPolice_precincts$agency), decreasing = TRUE)
```

## Redundant Data: Borough
```{r redundant_borough}
## 'borough' and 'park_borough' are identical
any(d311$borough != d311$park_borough)

## 'borough' and 'borough_boundaries' are almost identical
nonMatchingBorough <- subset(
    d311, borough_boundaries !=
          as.integer(str_replace_all(
              borough, c("STATEN ISLAND" = "1",  "BROOKLYN"      = "2",
                         "QUEENS"        = "3",  "MANHATTAN"     = "4",
                         "BRONX"         = "5",  "Unspecified"   = NA))),
    select = c("borough", "borough_boundaries", "agency"))
nrow(nonMatchingBorough)
```

## Redundant Data: Latitude, Longitude, and Location
```{r location}
print(head(d311[, c("latitude", "longitude", "location")]), digits = 16)
```

## Redundant Data: Cross Street and Intersection Street
```{r cross_intersection}
## one empty but the other not
nrow(subset(d311, cross_street_1 == "" & intersection_street_1 != ""))
nrow(subset(d311, cross_street_1 != "" & intersection_street_1 == ""))

## if both not empty, they are almost the same
head(subset(d311, cross_street_1 != "" & intersection_street_1 != "" &
                  cross_street_1 != intersection_street_1,
            select = c("cross_street_1", "intersection_street_1", "agency")))
```

##  Cross Street and Intersection Street (cont'd)
```{r extra_space}
## after removing extra spaces, most of them are the same
dupXStreets <- subset(
    d311, gsub(" ", "", cross_street_1) ==
          gsub(" ", "", intersection_street_1),
    select = c("cross_street_1", "intersection_street_1", "agency"))
nrow(dupXStreets)
head(dupXStreets)
```

##  Cross Street and Intersection Street (cont'd) 
```{r dup_streets}
table(dupXStreets$agency)
head(subset(dupXStreets, cross_street_1 != intersection_street_1))
```

## Less-Efficient Storage: Memory
```{r storage}
## code borough by integer
object.size(d311$borough_boundaries)
object.size(d311$borough)

## code agency by integer
object.size(d311$agency)
agency_code <- rep(seq(length(d311freq_agency)), d311freq_agency)
object.size(agency_code)
```

## Less-Efficient Storage: Disk
```{r storage-disk}
write.csv(subset(
    d311, select = -c(park_borough, borough_boundaries, location,
                      intersection_street_1, intersection_street_2,
                      agency_name)),
    quote=FALSE, row.names = FALSE,
    file = "../data/311_202303_smaller.csv")

(newsize <- file.size("../data/311_202303_smaller.csv"))
(oldsize <- file.size("../data/311_202303by202305.csv"))
(oldsize - newsize) / oldsize
```

# Conclusion

## Suggestions on Open Data Curation

+ Improve dictionary
    - Code categorical variables (e.g., agency) 
	- Code descriptive fields (e.g., complain type; vehicle type)
    - Make the descriptions informative (e.g., zip code)
+ Impose validity check
    - date (logical check)
    - legitimate values 
	- descriptive fields (drop-down menu, possibly multi-step)
    - missing value: coding; validity
+ Improve data storage/transportation efficiency
    - Eliminate redundant columns
    - Reduce storage with no information loss
+ Feedback: 
    - Open review for open data
	- Interaction with the maintenance team

# References {.allowframebreaks} 
\scriptsize