### 2023-06-10 talk/slides.Rmd
minor edits

---
title: "On Principles for Government Open Data Curation"
author: ""     # to be supplied through headers-includes below
institute: ""  # to be supplied through headers-includes below
output:
  beamer_presentation:
    theme: "AnnArbor"
    colortheme: "dolphin"
    slide_level: 2
#    toc: true
    number_sections: false
classoption: "aspectratio=169"
header-includes:
  - \usepackage{amsmath, amsfonts, bm}
  - \usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
  - \usepackage{caption}
  - \usepackage{booktabs}
  - \usepackage{multirow}
  - \usepackage{adjustbox}
  - \setbeamercolor{frametitle}{bg=white}
  - \AtBeginSection[]
    {
    \begin{frame}<beamer>
      \frametitle{Outline}
      \tableofcontents[currentsection]
    \end{frame}
    }
  - \AtBeginDocument
    {
    \title[NYC 311 Requests]{On Principles for Government Open Data Curation\\A Case Study with NYC 311 Servece Requests}
    \author[Jun Yan]{David Tussey\inst{1} and Jun Yan\inst{2}}
    \institute[University of Connecticut]{\inst{1} Former Executive Director, NYC DoITT\\
    \inst{2} University of Connecticut}
    }
bibliography: ["../manuscript/ref.bib"]
link-citations: true
link-color: true
# csl: apa.csl
---

```{r setup, include = FALSE, echo = FALSE}
library(stringr)
source("../code/functions.R")

def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  paste0("\n \\", "small","\n\n", x, "\n\n \\normalsize")
})
## Load the 311 SR data file
```

# Introduction

## NYC Open Data

- Open Data is viewed as a right for all [@janssen2012benefits; @wang2016adoption].
- Cities lead open data efforts wtih >50 cities with formal programs.
- NYC is a forerunner with Open Data Law in 2012 [@zuiderwijk2014open].
- Today NYC Open Data has 3511 datasets from 98 Agencies, accesible online 
  (\url{https://opendata.cityofnewyork.us}).
- Used by many researchers:
    + Health [@cantor2018facets, shankar2021data]
    + Smart city [@neves2020impacts]
    + Transportation [@gerte2019understanding]

## NYC 311 Service Requests

- 311 Service Request data is the most viewed dataset at NYC Open Data.
- NYC 311 is a unified frontline for all service requests to the NYC government agents.
- The data was used in STAT 3255/5255 Introduction to Data Science at UConn, with student presentations at 2022/2023 NYC Open Data Week.
- In this study, we exported 1-31 March 2023 311 SRs for analysis (~251K records).
- Data curation is critical for open data such as the 311 SRs [@witt2009constructing; @borgman2012conundrum; @mclure2014data; @barns2016mine; @hart2016ten; @beheshti2019datasynapse; @shankar2021data].
- Issues
    - Validity
	- Efficiency

# Issues with NYC 311 Requests Data

## Data Columns
<!-- - Column names: Remove trailing "."s.  Remove extra "."s -->
<!-- - Convert "." separator to "_".  Covert column names to lower cases -->
- Tidied column names (lower cases; underscore delimited; no trailing spaces)
- Last five columns not referenced in the Data Dictionary (not the case last year)
    + The NYC-OD Team advised these were "computed" fields derived from other data.

```{r read}
d311 <- read.csv("../data/311_202303by202305.csv", header = TRUE)
names(d311) <- makeColNamesUserFriendly(names(d311))
dim(d311)
tail(names(d311))
```

## Contributions by Agencies

```{r byAgency}
d311freq_agency <- table(d311$agency)
sort(d311freq_agency, decreasing = TRUE)
```

## Missing data: Character columns
- Various forms: "", "unspecified", and "unknown"
- Results show 3 groupings: all blanks (>90%), a few blanks (12-44%), or nearly no blanks (1-5%)

```{r blanks}
cMissingDataPerColumn <- charColumnsMissingData(d311)
cMissingDataPerColumn[order(-rowSums(cMissingDataPerColumn)), ]
```

## Missing data: Numeric Colummns

```{r missing}
nMissingDataPerColumn <- numColumnsMissingData(d311)
sort(nMissingDataPerColumn, decreasing = TRUE)
```

<!-- ## Unspecified/Unknown data columns -->
<!-- - Several columns have many "Unspecified" or "Unknown" values. These are deemed valid values. -->
<!-- ```{r uknown} -->
<!-- tail(missingDataPerColumn[ order( -missingDataPerColumn[ , 2] ), ], 4) -->
<!-- ``` -->

## Invalid Data: Created/Closed Date
```{r date}
t1 <- strptime(d311$closed_date, format = '%m/%d/%Y %I:%M:%S %p')
t0 <- strptime(d311$created_date, format = '%m/%d/%Y %I:%M:%S %p')
tt <- as.numeric(difftime(t1, t0, units =  "day"))
badTime <- subset(d311, tt <= 0,
                  select = c("created_date", "closed_date", "agency"))

## exactly the same created and closed date down to the second; a lot!
sum(tt == 0, na.rm = TRUE)
table(subset(badTime, created_date == closed_date, select = "agency"))
```

## Bad Closed Date
```{r baddate}
## quite some negative response times
table(subset(badTime, created_date != closed_date, select = "agency"))

summary(tt[tt < 0 & (!is.na(tt))])
```

## Invalid Data: Incident Zipcodes
```{r zip1}
validZipcodes <- scan("../data/NYCzipcodes.csv")

badIncident_zip <- subset(d311, ! incident_zip %in% validZipcodes,
                          select = c("incident_zip", "agency"))
head(sort(table(badIncident_zip$incident_zip), decreasing = TRUE))
```

## Invalid Data: Zipcodes (One of the last 5 columns)
```{r zip2}
badZipcode  <- subset(d311, ! zip_codes %in% validZipcodes,
                      select = c("zip_codes", "agency"))
nrow(badZipcode)  # mostly invalid! One possibility: debt collectors
```

## Invalid Data: Borough
```{r borough}
validBorough <- c("BRONX", "BROOKLYN", "MANHATTAN", "QUEENS", "STATEN ISLAND",
                  "Unspecified")
badBorough <- subset(d311, ! borough %in% validBorough,
                     select = c("borough", "agency"))
head(table(badBorough$agency))
head(table(subset(d311, borough == "Unspecified")$agency))
```

## Invalid Data: Borough_boundaries and Park_borough
```{r borough2}
validBB <- c("1", "2", "3", "4", "5")
badBB <- subset(d311, ! borough_boundaries %in% validBB,
                select = c("borough_boundaries", "agency"))
head(table(badBB$agency))

badPark_borough <- subset(d311, ! park_borough %in% validBorough,
                          select = c("park_borough", "agency"))
head(table(badPark_borough$agency))
```

## Invalid Data: Open Data Channel
"Due to pandemic call handling modifications, the ‘Open Data Channel Type’ values may not accurately indicate the channel the Service Request was submitted in for the period starting March 2020."

```{r channel}
validChannel <- c("MOBILE", "ONLINE", "PHONE", "OTHER")
badChannel_type <- subset(d311, ! open_data_channel_type %in% validChannel,
                          select = c("open_data_channel_type", "agency"))
head(table(badChannel_type$agency)) ## UNKNOWN
```

## Invalid Data: Police Precincts
```{r nypd}
validPolice_precincts <- scan("../data/NYPDPrecincts2023.csv", skip = 1)
badPolice_precincts <- subset(
    d311, ! police_precincts %in% validPolice_precincts,
    select = c("police_precincts", "agency"))
sort(table(badPolice_precincts$agency), decreasing = TRUE)
```

## Redundant Data: Borough
```{r redundant_borough}
## 'borough' and 'park_borough' are identical
any(d311$borough != d311$park_borough)

## 'borough' and 'borough_boundaries' are almost identical
nonMatchingBorough <- subset(
    d311, borough_boundaries !=
          as.integer(str_replace_all(
              borough, c("STATEN ISLAND" = "1",  "BROOKLYN"      = "2",
                         "QUEENS"        = "3",  "MANHATTAN"     = "4",
                         "BRONX"         = "5",  "Unspecified"   = NA))),
    select = c("borough", "borough_boundaries", "agency"))
nrow(nonMatchingBorough)
```

## Redundant Data: Latitude, Longitude, and Location
```{r location}
print(head(d311[, c("latitude", "longitude", "location")]), digits = 16)
```

## Redundant Data: Agence, Agency Name, and Other Categorical Variables
```{r agent}
head(d311[, c("agency", "agency_name")])
head(sort(table(d311$complaint_type), decreasing = TRUE)) # 174 types
```

## Redundant Data: Cross Street and Intersection Street
```{r cross_intersection}
## one empty but the other not
nrow(subset(d311, cross_street_1 == "" & intersection_street_1 != ""))
nrow(subset(d311, cross_street_1 != "" & intersection_street_1 == ""))

## if both not empty, they are almost the same
head(subset(d311, cross_street_1 != "" & intersection_street_1 != "" &
                  cross_street_1 != intersection_street_1,
            select = c("cross_street_1", "intersection_street_1", "agency")))
```

##  Cross Street and Intersection Street (cont'd)
```{r extra_space}
## after removing extra spaces, most of them are the same
dupXStreets <- subset(
    d311, gsub(" ", "", cross_street_1) ==
          gsub(" ", "", intersection_street_1),
    select = c("cross_street_1", "intersection_street_1", "agency"))
nrow(dupXStreets)
head(dupXStreets)
```

##  Cross Street and Intersection Street (cont'd) 
```{r dup_streets}
table(dupXStreets$agency)
head(subset(dupXStreets, cross_street_1 != intersection_street_1))
```

## Less-Efficient Storage: Memory
```{r storage}
## code borough by integer
object.size(d311$borough_boundaries)
object.size(d311$borough)

## code agency by integer
object.size(d311$agency)
agency_code <- rep(seq(length(d311freq_agency)), d311freq_agency)
object.size(agency_code)
```

## Less-Efficient Storage: Disk
```{r storage-disk}
write.csv(subset(
    d311, select = -c(park_borough, borough_boundaries, location,
                      intersection_street_1, intersection_street_2,
                      agency_name)),
    quote=FALSE, row.names = FALSE,
    file = "../data/311_202303_smaller.csv")

(newsize <- file.size("../data/311_202303_smaller.csv"))
(oldsize <- file.size("../data/311_202303by202305.csv"))
(oldsize - newsize) / oldsize
```

# Conclusion

## Suggestions on Open Data Curation

+ Improve dictionary
    - Code categorical variables (e.g., agency) 
	- Code descriptive fields (e.g., complain type; vehicle type)
    - Make the descriptions informative (e.g., zip code)
+ Impose validity check
    - date (logical check)
    - legitimate values 
	- descriptive fields (drop-down menu, possibly multi-step)
    - missing value: coding; validity
+ Improve data storage/transportation efficiency
    - Eliminate redundant columns
    - Reduce storage with no information loss

## Outlook

- Open review for open data

- Interaction with the maintenance team

- Open data curation protocol

- Open data analytic portal

# References {.allowframebreaks} 
\scriptsize