### 2024-05-21 manuscript/311 manuscript.tex
Console output for latest code revisions

\documentclass[12pt, titlepage]{article}

\usepackage[margin = 1in]{geometry}
\usepackage{amsmath}
\allowdisplaybreaks
\usepackage{authblk}
\usepackage{setspace}
\usepackage{natbib}
\usepackage{hyperref}\hypersetup{colorlinks=true, citecolor=blue}
\usepackage{parskip}
\usepackage{textgreek}
\usepackage{graphicx}
\usepackage{booktabs} % For better looking tables
\usepackage{siunitx} % For alignment of numbers
\usepackage{placeins}
\usepackage{caption}  % For custom captions

\usepackage[]{lineno}
\linenumbers*[1]
%% patches to make lineno work better with amsmath
\newcommand*\patchAmsMathEnvironmentForLineno[1]{%
	\expandafter\let\csname old#1\expandafter\endcsname\csname
	#1\endcsname
	\expandafter\let\csname oldend#1\expandafter\endcsname\csname
	end#1\endcsname
	\renewenvironment{#1}%
	{\linenomath\csname old#1\endcsname}%
	{\csname oldend#1\endcsname\endlinenomath}}%
\newcommand*\patchBothAmsMathEnvironmentsForLineno[1]{%
	\patchAmsMathEnvironmentForLineno{#1}%
	\patchAmsMathEnvironmentForLineno{#1*}}%
\AtBeginDocument{%
	\patchBothAmsMathEnvironmentsForLineno{equation}%
	\patchBothAmsMathEnvironmentsForLineno{align}%
	\patchBothAmsMathEnvironmentsForLineno{flalign}%
	\patchBothAmsMathEnvironmentsForLineno{alignat}%
	\patchBothAmsMathEnvironmentsForLineno{gather}%
	\patchBothAmsMathEnvironmentsForLineno{multline}%
}

\sisetup{
  group-separator = {,},
  round-mode = places,
  round-precision = 0
}

% control floats
\renewcommand\floatpagefraction{.9}
\renewcommand\topfraction{.9}
\renewcommand\bottomfraction{.9}
\renewcommand\textfraction{.2}
\setcounter{totalnumber}{50}
\setcounter{topnumber}{50}
\setcounter{bottomnumber}{50}

\newcommand{\jy}[1]{\textcolor{orange}{JY: (#1)}}
\newcommand{\yh}[1]{\textcolor{purple}{YH: (#1)}}
\newcommand{\yx}[1]{\textcolor{cyan}{YX: (#1)}}
\let\proglang=\textsf
%% \newcommand{\pkg}[1]{{\fontseries{m}\selectfont #1}}
%% \newcommand\code[2][black]{\textcolor{#1}{\texttt{#2}}}

\title{Principles for Open Data Curation: A Case Study with the New
  York City 311 Service Request Data}

\author[1]{David Tussey}
\author[2]{Jun Yan}
\affil[1]{Former Executive Director, NYC DoITT}
\affil[2]{Department of Statistics, University of Connecticut}

\begin{document}
\maketitle

\tableofcontents % Optional: Table of Contents
\listoffigures % List of Figures
\listoftables % List of Tables

\begin{abstract}
  Open data is important ...
  Quality control is important ...
  We use New York City 311 requests data  ...
  We suggest some open data quality control protocols ...

\bigskip

\noindent
{\it Keywords:}
consistency check;    
data release protocol;
data science;
open data;
data cleansing;
quality control	
\end{abstract}

\doublespacing

\section{Introduction} \label{sec:intro}

In the early 21st century, the open data movement began to take shape, driven by the 
fundamental belief that freely accessible data can transform 
both societies and governments. This movement champions the principles of 
transparency, innovation, and public engagement.

A landmark in this journey was the launch of the United States'
\href{https://www.data.gov}{Data.gov} portal in 2009, a pioneering
platform in making government data widely accessible. Shortly after,
the European Union followed suit, unveiling its
\href{https://data.europa.eu/euodp}{Open Data Portal} in 2012, further
cementing the movement's global reach. Furthermore, the World Bank's Open
Data initiative, initiated in 2010, stands out as a comprehensive
repository for global development data, available at
\href{https://data.worldbank.org}{World Bank Open Data}.

These initiatives represent significant strides in democratizing data, in breaking barriers that once
kept valuable information on government performance in silos. Their collective impact is profound, extending beyond
mere data sharing to fostering a culture of openness that benefits
individuals, communities, governments, and economies worldwide
\citep{barns2016mine, wang2016adoption}.

New York City (NYC) has emerged as a forerunner in the open data
movement, marked by the enactment of the Open Data Law in 2012
\citep{zuiderwijk2014open}. This landmark legislation led to the
creation of the \href{https://opendata.cityofnewyork.us}{NYC Open Data
  portal}, which today hosts an impressive array of 2.700 datasets
from 80 different city agencies. This resource has become invaluable
for researchers in various fields as well an enabling a high degree of local government transparency

In health, datasets have enabled
significant studies on facets of health and healthcare delivery
\citep{cantor2018facets, shankar2021data}. In the realm of urban
development, data has been instrumental in advancing smart city
initiatives \citep{neves2020impacts}. Additionally, transportation
research has benefited greatly from this wealth of data, aiding in the
understanding of urban mobility and infrastructure
\citep{gerte2019understanding}.

NYC's Open Data initiative not only
exemplifies commitment to transparency and public engagement but also
illustrates how open data can be a powerful tool in addressing complex
urban challenges.

Data curation is fundamental in the open data ecosystem, ensuring the
utility and reliability of datasets for diverse applications. Among
the earliest discussions, \citet{witt2009constructing} focus on the
development of data curation profiles tailored to specific contexts,
setting a precedent for targeted data management
strategies. Addressing broader challenges in data sharing and
management, \citet{borgman2012conundrum} highlights the complexities
of research data distribution, emphasizing the need for robust
strategies. This is complemented by the work of \citet{hart2016ten},
who outline essential principles for effective data management,
particularly emphasizing the importance of meticulous curation
practices.

In the realm of collaborative data management,
\citet{beheshti2019datasynapse} underscore the significance of
cooperative environments for managing and sharing social data
effectively. This aspect of data curation gains further relevance in
the research by \citet{mclure2014data}, which delves into the specific
practices and needs within data curation communities. The practical
implications of data curation are vividly illustrated in the context
of public health and global challenges. \citet{cantor2018facets}
demonstrate the utility of curated open data in evaluating community
health determinants. Furthermore, the COVID-19 pandemic serves as a
real-world example, with \citet{shankar2021data} observing the
critical role of collective data curation efforts in managing and
responding to the crisis.

Collectively, these studies not only highlight the multifaceted nature of data curation but also emphasize
its indispensable role in enhancing the applicability and value of
open data across various domains.

The contributions of this paper are twofold. First, we delve into
the specifics of data curation challenges using the NYC 311 Service
Request Data as a case study. This renowned and frequently viewed 
dataset serves as a prime example for examining key issues in data curation, 
including data validity, consistency, and curation efficiency. 
We illustrate these points with live examples drawn from our processing 
of the 311 data.

Secondly, building upon insights gained from this case study, we 
propose a set of data curation principles tailored for government-released open data. 
These principles are designed to address the unique challenges 
and requirements observed in the curation of such datasets.

The remainder of this paper is organized as follows. 
Section~\ref{sec:data} offers a brie review of the 
history and current status of the 311 requests data. In 
Section~\ref{sec:issues}, we identify and list potential issues 
impacting data quality and curation efficiency. 
Section~\ref{sec:improve} is dedicated to providing 
actionable suggestions for mitigating or resolving these identified issue. 
Following this, Section~\ref{sec:protocol} outlines a series of general 
principles for the release of open data, drawing from our findings. 
The paper concludes with a discussion in Section~\ref{sec:disc}, encapsulating 
the key insights and implications of our research.

\section{NYC 311 Service Requests Data} \label{sec:data}

The NYC 311 service, a critical component of New York City's public
engagement and service response framework, serves as a centralized hub
for non-emergency inquiries and requests. Introduced in 2003, the NYC
311 system was designed to streamline the city's response to
non-emergency issues, ranging from noise complaints to street
maintenance requests. The evolution of this system can be traced from
its initial implementation as a simple inquiry channel to a
comprehensive data management system that handles millions of requests
annually. Key milestones in its development include:

\begin{itemize}
 	\item 2003 - NYC 311 system  goes live with phone-based call center only.
   	\item 2009 - 311 online \& mobile apps launched. Today 60\% of all Service Requests come from mobile \& online channels.
   	\item 2019 - Major software upgrade of 311 system with enhanced capabilities and cloud-based resilience. 	
      	\item 2020 - In August, driven by the COVID pandemic, monthly 311 Service Requests hit an all-time high of 348,463. Seven of the Top 10 busiest days occur in 2020.
      	\item 2021 - 311 System is expanded to into the MTA's city subway system, the largest expansion since its inception.
   	\item 2023 - Record high year for 311 with 3.23 million Service Requests.
\end{itemize}

Today, the NYC 311 data system is a robust platform that manages a
vast array of urban living-related inquiries. As of 2024, the system handles approximately 3.5 million service 
requests per year, covering categories such as noise complaints, illegal parking, heat/hot water complaints, abandoned vehicles,
and unsanitary conditions. The data, managed through a tailored application built on the Microsoft Dynamics platform, is a valuable resource for city
administration and policy-making. It is publicly accessible through the 
\href{https://opendata.cityofnewyork.us/}{NYC Open Data Portal}. The online Portal provides access
to all the data sets (approx. 2700) and provides a convenient method for querying, grouping, aggregating, geo-mapping, visualization, and exporting
results. This open data initiative enables not only governmental
transparency but also empowers researchers, civic developers, and the
general public. Examples of its use include:

\begin{itemize}
	\item NYC restaurant violations
	\item Popular baby names
	\item Mapping of car crashes involving pedestirans
	\item Mapping of sidewalk widths in NYC 
	\item Visualization of NYC High School and College enrollment
	\item Interactive dashboard to filter and view charges for jail inmates
	\item Internet access points.
	\item ...and 220 different complaint types. 
\end{itemize}

Additional Open Data projects can be found at \href{https://opendata.cityofnewyork.us/projects/}{NYC Open Data Project Gallery}
and at \href{https://beta.nyc/beta/products/}{\textbeta etaNYC} products and tools.

Despite its success, the system faces many challenges such as:
\begin{itemize}
	\item Data timeliness, accuracy, and consistency
	\item Difficulties correlating data over long time periods, e.g. over a 10-year period
	\item Data anonymization and handling Personally identifiable information (PII)
	\item Integration with stand-alone systems at selected NYC Agencies
	\item Managing API usage, authentication, and load for 3rd party users
	\item Incorporating ongoing technology changes and upgrades 
	\item Understanding the role of chatbots and AI to provide online customer service
\end{itemize}

These (and other) challenges are constantly being addressed by the various agency open data managers
as well as the \href{https://www.nyc.gov/content/oti/pages/}{NYC Office of Technology and Innovation} which
provides the technology support for the open data system.  
Many of the underlying datasets are updated daily, usually running 1-2 days behind from creation.
However, a number of datasets are updates monthly (such as restaurant inspections), and some are
not updated for one or more years (such as community board boundaries). This poses problems when 
merging these dataset with one another, as would be expected. Typically, the data must be
exported and manipulated in another program

Data accuracy and consistency is typically established between OTI and the other contributing
City Agencies, such as NYPD, Housing Preservation and Development (HPD), Parks \& Recreation, etc.,
via Service Level Agreements (SLAs). However, enforcement of those SLAs is often challenging. The 
result is that often data accuracy and logical consistency is occasionally compromised.

The NYC 311 organization, part of OTI, operates a highly tailored and sophisticated case
management system based, as noted, on Microsoft Dynamics. However, when the original 311
system was developed in 2003, a decision was made to allow select NYC Agencies to
maintain their own local computer systems, such as the mainframe systems at the NYC Department
of Sanitation (DSNY) and others. And when the 311 system was completely upgraded in 2019, those
same integration were `honored` in order to reduce the scope of the project. (Perhaps in
retrospect, while a good budgetary decision, perhaps not such a good technology decision.)

The result was that at least nine NYC Agencies have local systems that are integrated either via
an API or custom code. These Agencies include:

\begin{itemize}
	\item Department of Social Services (DSS), one of the largest City Agencies
	\item Department of Sanitation, although recently part of DSNY has moved to the core 311 software.
	\item Department of Parks \& Recreation (forestry complaints only)
	\item Housing, Preservation \& Development (fully integrated)
	\item Department of Environmental Protection (DEP) (fully integrated)
	\item Department of Healthy \& Mental Hygiene (fully integrated) 
	\item Department of Consumer \& Worker Protection (DCWP - formerly known as the Dept of Consumer Affairs)
	\item New York Police Department (NYPD) (integrated with their reporting system)
	\item Department of Business (DoB) (fully integrated)
\end{itemize}

One result of these integrations, and other Agencies that use the 311 APIs, is that some of the data accuracies
and data consistency checks inherent in the 311 system are not applied uniformly, and in some cases not at all.
Errors often creep in from other Agency systems and are not easily corrected in the 311 application.

One area that will almost certainly remain an ongoing challenge is the incorporation of new software, hardware,
and cloud services which are constantly emerging.  Telephonic integration recently was reconfigured due to a new
telephone vendor. Desktop software at the 311 Agency are subject to the same upgrades that home users feel. And while the 
311 system was designed to be easily configured with new features, some emerging business issues as well as
technologies simply require additional software engineering efforts to accommodate.

One area in particular which is heavily influencing customer outreach for the 311 system is the use of
online chatbots. Increasing these are being influenced by Artificial Intelligence (AI) advances. It is a given that
a number of citizen's interactions with City government could be handled accurately and quickly by AI-driven
applications. How to do that while preserving customer privacy and anonymization are challenging issues.

The impact of the NYC 311 data extends beyond operational efficiency;
it has become instrumental in shaping City governance and community
engagement. The data has been pivotal in such areas as:

\begin{itemize}
	\item Advice on shelters during hurricanes and winter blizzards
	\item Countless inquires during the COVID pandemic (testing centers, vaccine sites, mobile vaccine clinics, etc.)
	\item Enforcement of standards between landlords and tenants.
	\item Re-allocation of NYC's yellow taxi routes based upon an insightful analysis conducted by the Taxi \& Limousine Commission (TLC)
	\item Improved responsiveness for City Agencies with direct client-facing responsibilities such as Parks \& Rec, Dept of Transportation, DSNY, etc.
	\item Increased 311 SRs related to Homelessness resulted in the launch of the City's Homeless Outreach \& Mobile Engagement Street Action Teams (HOME-STAT)
	which includes street canvassers and increases in homeless housing facilities.
	\item A significant increase in graffiti complains in 2017 led directly to the `Graffiti-Free NYC'' campaigns and increased funding for removal of graffiti in public spaces
	\item Enhanced geospatial data regarding the location of incidents, the responding Agency, and actions taken.
	\item Establishment of Agency SLAs in regard to certain complaint types, such as residential noise, homelessness person assistance, street light outages, 
	rat sightings, and other neighborhood quality of life issues.
\end{itemize}

The NYC 311 service exemplifies the dynamic nature of urban data
management and its critical role in modern governance. It provides an
invaluable resource for data scientists to practice data science
activities, including data curation.

We extracted  311 Service Request (SR) data from the period of CY 2022-2023, a dataset consisting of
6.4 million records. This dataset serves as the primary source for the analysis effort. This is a large dataset which enables 
observation of some rarely seen events, data inaccuracies, and logical inconsistencies. It provides enough data to examine trends and data outlier analysis. 
We query the data directly from the NYC Open Data Portal and then export the data in CSV format for analysis in a custom written R programs.

Additionally, we extracted a 10-year dataset consisting of the years 2014-2023; some 27 million rows. Albeit with just a few columns, 
this very large dataset allows for long-term trend observations and confirms that the analysis using the 
2-year dataset used for the majority of the analysis.

\section{10-year Analysis \& Trends} \label{sec:history}

Since the NYC 311 System went live in 2003, the usage has grown at an annual
rate of 4.6\%. For this study, we also looked at a 10-year period (2014-2023) to
observe certain system performance characteristics over a longer time-frame . 
During that time-frame, the number of 311 SRs grew 50\% as shown in the chart below which shows the yearly SR counts.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{Yearly.png}
  \caption{SR counts by year for a 10-year period}
  \label{fig:yearly-counts}
\end{figure}

This 10-year timeframe (2014-2023) included the COVID pandemic impact. Note that the pandemic
impact reached a peak in the July-September 2020 timeframe, accompanied by additional spikes
in the 2\textsuperscript{nd}half of 2021 as well. The all-time daily high for SRs occurred on Tuesday, 2020-08-04, with a spike of
24,415 SRs, a full 9.8 \textsigma's from the mean.

This next chart shows the 10-year period by calendar month. Note that almost all the months before the COVID pandemic outbreak in Feb/Mar
of 2020 are ``below average'', while almost all the months after that period are ``above average'' months. And that trend of much higher usage
of the 311 system has continued into the early months of 2024.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{Monthly.png}
  \caption{SR counts by month for a 10-year period}
  \label{fig:monthly-counts}
\end{figure}

The daily tend over that same 10-year period is also interesting, as it shows both significant spikes
and significant dips; a noisy graph. While the spikes in 2020 and 2021 appear COVID related, we are unable to 
ascertain any reason for the significant spike on Saturday, 2023-09-02 where 115,205 SRs were created.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.3]{Daily.png}
  \caption{SR counts by day for a 10-year period}
  \label{fig:daily-counts}
\end{figure}

Seasonal trends are also prevalent. The months of November-thru-April are ``below average'' months, while the months May-thru-October are ``above average''.
 This seasonal trend is observed throughout the 10-year period.

\begin{figure}[htbp]
  \centering
  \includegraphics[scale=0.65]{Calendar-Month.png}
  \caption{SR counts by calendar month for a 10-year period}
  \label{fig:calendar-months-counts}
\end{figure}

Of note, the lowest SR counts-per-day occur on Christmas  period (+/- a day), in 5 of the Bottom 10 SR counts-per-day. The summer moths and early
fall months make up all of the Top 10 SR counts-per-day. As previously noted, the all-time daily high for SRs occurred on Tuesday, 2020-08-04 with the 10-year
low number occurs on Christmas Day in 2016, a Sunday.

\begin{table}[htbp]
	\begin{minipage}[t]{0.35\textwidth} % Adjust the width of the minipage as needed
	\centering
	\small % Adjust the font size
	\caption{Top 10 days}
	\label{tab:top_ten_counts}
		\begin{tabular}{@{}lll@{}}
			\toprule
			Date & Count & Rank \\
			\midrule
			2020-08-04 & 24415 & 1 \\
			2020-08-05 & 19560 & 2 \\
			2023-09-29 & 17962 & 3 \\
			2020-07-05 & 16916 & 4 \\
			2020-06-21 & 15883 & 5 \\
			2020-06-20 & 15825 & 6 \\
			2020-07-04 & 15794 & 7 \\
			2021-09-02 & 15205 & 8 \\
			2020-06-28 & 14057 & 9 \\
			2021-10-28 & 13575 & 10 \\
			\bottomrule
		\end{tabular}
	\end{minipage}\hfill % Adjust the space between the tables
	\begin{minipage}[t]{0.35\textwidth} % Adjust the width of the minipage as needed
	\centering
	\small % Adjust the font size
	\caption{Bottom 10 days}
	\label{tab:bottom_ten_counts}
		\begin{tabular}{@{}lll@{}}
			\toprule
			Date & Count & Rank \\
			\midrule
			2014-12-25 & 2965 & 1 \\
			2015-12-25 & 3247 & 2 \\
			2014-04-20 & 3287 & 3 \\
			2015-12-26 & 3405 & 4 \\
			2016-11-06 & 3419 & 5 \\
			2014-07-04 & 3455 & 6 \\
			2014-12-27 & 3534 & 7 \\
			2014-09-14 & 3563 & 8 \\
			2014-02-02 & 3565 & 9 \\
			2016-12-25 & 3587 & 10 \\
			\bottomrule\\
		\end{tabular}
	\end{minipage}
\end{table}

And although the highest number of SRs occur in July and the lowest in February,
when the \#-of-SRs-per-day metric is computed by adjusting for the differing number of days for each month -- the
busiest month is shown to be June (followed by July and September). And similarly the quietest months are April, March, and 
February. In general, the winter months see fewer SRs while the summer months show 
an increase in activity, as previously noted.

\begin{table}[ht]
    \centering
    \begin{tabular}{lS[table-format=7.0]S[table-format=6.0]} % 'l' for left-aligned, 'S' for siunitx number alignment
        \toprule
        \textbf{Month}     & \textbf{Count}   & \textbf{Count per Day} \\ 
        \midrule
        June  & 2454156 & 81805 \\ 
        July  & 2461525 & 79404 \\ 
        September & 2352822 & 78427 \\ 
        October & 2407238 & 77653 \\ 
        August & 2401731 & 77475 \\ 
        May  & 2334466 & 75305 \\ 
        November  & 2236037 & 74535 \\ 
        January   & 2271524 & 73275 \\ 
        December  & 2199765 & 70960 \\ 
        February  & 1996408 & 70795 \\ 
        March & 2174349 & 70140 \\ 
        April  & 2058169 & 68606 \\ 
       \bottomrule
    \end{tabular}
    \caption{Monthly counts sorted by counts-per-day}
    \label{tab:monthly_counts}
\end{table}

\FloatBarrier % Ensure all floats before this point are processed before moving on

Two  areas that are important in this data cleansing analysis effort are the responsible NYC Agency, and the actual type of complaint.
Given both the size of the City of New York government, it's typically necessary to identify the responsible Agency as well as the type of complaint in order to 
clarify the responsible area and assist in troubleshooting any discrepancy.

In discovering a data error, we typically observe two trends; either a single (or a small few) Agencies are responsible (such as Dept of Transportation), or it is a system-wide issue that foolows the general distribution
of  Service Requests (SRs) across the system. Let's take a brief look at how this plays out across the system, so when we identify a
issue, it will be useful to see how that issue flows back to a specific Agency, and thus establish responsibility.

Although the 311 SRs involve 16 different agencies, there are the ``big six'' Agencies that comprise
90\% of the complaints. These six are (in order):

\begin{itemize}
	\item New York Police Department (NYPD)
	\item Housing Preservation \& Development (HPD
	\item New York City Department of Sanitation (DSNY)
	\item Department of Transportation (DOT)
	\item Department of Environmental Protection (DEP)
	\item Department of Parks \& Recreation (DPR)
\end{itemize}

\begin{figure}[htbp]
  \centering
	  \includegraphics[scale=0.65]{SRs_by_Agency.png}
	  \caption{SR counts by Agency and Cumulative Percentage}
	  \label{fig:SR_counts_by_Agency}
\end{figure}

It is also interesting to note the distribution of the type of complaints in the 311 SR dataset; it is heavily skewed to  a select few complaints.

\begin{figure}[htbp]
  \centering
	  \includegraphics[scale=0.65]{SRs_by_Complaint_Type.png}
	  \caption{Top 20 SR Complaints and Cumulative Percentage}
	  \label{fig:SR_complaints}
\end{figure}

The top 20 complaints contain several of the eight different types of ``Noise'' complaints (Residential, Commercial, Street, Helicopter, etc.). These Noise-related
complaints total 1.435 million in the 2022-2023 dataset, a full 22\% of all complaint types, which is the most frequently occurring type. (Note: Ilegal Parking is 
second, followed by Heat/Hot Water.) Noise complaints are also handled by NYPD, as is Illegal Parking and Blocked Driveway complaints, which mostly
accounts for the top ranking of 311 SRs being routed to the NYPD.

The NYC HPD handles about 49\% of what NYPD handles. For  non-NYC readers, 
HPD is the City agency that manages the  177,569 NYC public housing units as well as monitoring all rental and leased apartment units. HPD collectively
serves 528,105 people, a population larger than Atlanta or Miami. Hence the large number of Heat/Hot Water and Unsanitary complaints.

\section{Potential Issues} \label{sec:issues}

What is data cleansing?  Wikipedia \href{https://en.wikipedia.org/wiki/Data_cleansing}{Data Cleansing} offers a good definition.

``Data cleansing or data cleaning is the process of detecting and correcting (or removing) corrupt or inaccurate records from a record set, 
table, or database and refers to identifying incomplete, incorrect, inaccurate or irrelevant parts of the data 
and then replacing, modifying, or deleting the dirty or coarse data.''

Many quality criteria are required in order to process high-quality data. These include:

\begin{itemize}
	\item Data validation - this effort can span a number of criteria
	\begin{itemize}
		\item Mandatory fields: Certain data fields cannot be empty
		\item Data-types: Certain fields must be of the correct type, e.g. numeric, character, date, 5 numeric digits, etc. Typically these are referenced in a Data Dictionary.
		\item Domain adherence:  Many data fields must adhere to a specific domain of values, e.g. statuses, state names, zip codes, gender. This includes data
		fields that are restricted to certain ranges, such as longitudes and latitudes.
	\end{itemize}   
	\item Structural errors to include naming conventions, extra fields no in the Data Dictionary, inconsistent data entry.
	For example intermixing blanks, space, NA, N/A. and <NA> to indicate the absence of data.
	\item Redundant, unnecessary, or inappropriate fields
	\item Logical inconsistencies such as two related fields that violate the nature of that relationship, such as a ``due date'' that is before the ``created date''.
	\item Data standardization. Often free-form entry fields can suffer from inconsistent data structures. Efforts to standardize these fields can greatly assist analysis.
	\item Accuracy and precision, which of course are not the same thing 
\end{itemize}

The effort of this analysis effort is to identify the presence of such errors; not to correct them. That effort would be undertaken after an investigation as to the why
and how such errors came about. Often this requires discussions with subject matter experts and liaison with various NYC Agency open data coordinators
to identify the original source of the error. For example, an invalid date field that is provided to the NYC Open Data data lake via an Agency integration could
be the result of selecting the wrong field from the source Agency's system, e.g. due date instead of closed date. In other cases, without a subject matter
expert, it is not possible to actually determine if the data is inaccurate or not. The R code used to analyze the 311 Service Request dataset identifies
both obvious inaccuracies as well as potential errors requiring further investigation.

This paper will focus on a set of specific data cleansing areas:

\begin{itemize}
	\item Structural issues with the data. Is it in the expected format?
	\item Missing/blank data
	\item Data type correctness
	\item Invalid values
	\item Logical inconsistencies and inconsistent patterns in the data
	\item Accuracy and precision
	\item Redundant data  
\end{itemize}

\subsection{Structural Issues}

Structural issues in the context of data cleansing refers to issues related to the how data is organized, formatted, or structured within a dataset. Structural issues 
can make it difficult to analyze the data effectively. Some common structural concerns in this 311 SR 2022-2023 dataset include:

\begin{itemize}
	\item Data structure (data fields, columns, formats, data types, etc.) not corresponding to the Data Dictionary
	\item Verification of correct data types (numeric, character, geospactial, dates, etc.)
	\item Embedded or combined values: data elements that contain multiple pieces of information 
\end{itemize}

In this dataset the writers discovered several structural issues.

	\subsubsection{Fields not in the Data Dictionary}
	The 311 SR \href{https://data.cityofnewyork.us/api/views/erm2-nwe9/files/b372b884-f86a-453b-ba16-1fe06ce9d212?download=true&filename=311_ServiceRequest_2010-Present_DataDictionary_Updated_2023.xlsx}{Data Dictionary} identifies 41 data columns (fields) along with some
	related information for each column. However, when you either download the data or explore it via the online portal,
	you will immediately note that there are 47 fields. The additional six fields are not addressed in the Data Dictionary, and show up on the Column Manager
	page as ``@computed\_region\_xxxx\_xxxx''. To some extent one can infer what the these @computed\_fields are, but it is not possible to know
	for certain. Here is a screenshot of the NYC Open Data Portal ``Column Manager'' page showing those columns:

	\begin{figure}[htbp]
	  \centering
		  \includegraphics[scale=0.65]{computed_columns_screenshot.png}
		  \caption{Screenshot of data portal showing the ``@computed'' columns}
		  \label{fig:computed-columns}
	\end{figure}

	This, of course, raises several questions:
	\begin{itemize}
		\item How are these fields computed? Using what source data? Using what computational method?
		\item How can these fields be verified or cross-checked?
		\item Why are their two fields that appear redundant ``Police Precinct'' and ``Police Precincts''?
		\item What are ``borough boundaries''? Are these the regular NYC boroughs?
		\item What are ``community districts''? Is this the same as the ``Community Boards'' that are part of the NYC government?
		\item Can these fields be included in the official 311 SR Data Dictionary?
		\item Can these fields be used for analytical purposes? 
	\end{itemize}

	The last bullet point is the key here. Are these fields reliable and accurate, and can they be used for subsequent analysis? Unfortunately,
	as of this date, those questions have not been fully addressed by the  NYC Open Data Team. This paper will address further issues with the
	accuracy and subsequent analytical usage for these six @computed fields.

	\subsubsection{Verifying data types}
	Fortunately, both the Data Dictionary and the portal column manager do a good job of indicating the data type.
	As indicated by the red underline in Figure \ref{fig:computed-columns}, there is a small icon next to the column name 
	which indicates the data type of the column. The stylized capital ``T'' indicates text, the ``\#'' sign indicates 
	a numeric, a calendar icon indicates a date field, and the ``pin on a map'' symbol indicates a geospatial field.

	The following fields were checked for their specified data types:
		\begin{itemize}
		\item created\_date, closed\_date, due\_date, and resolution\_action\_updated\_date are all in proper date format.
		\item zip\_codes and incident\_zip: All are 5 numeric digits except for 2 non-numeric entries for incident\_zip (``na'' and ``N/A'').
		\item x\_coordinate\_state\_plane \&  y\_coordinate\_state\_plane (a NY State geo-location system) are all numeric.
		\item latitude \& longitude are both numeric.
		\item community\_districts, borough\_boundaries, city\_council\_district, police\_precinct, and police\_precincts are all numeric.
		\item Free-form text fields that contain a ``,'' (e.g. incident\_address, resolution\_description) are enclosed in quotation marks.
		\end{itemize}

\subsection{Missing-Blank data}

\subsection{Data type correctness}

\subsection{Invalid values}

\subsection{Logical inconsistencies }

\subsection{Accuracy and precision}

\subsection{Redundant data}

\subsection{Case Study I - Response to Requests for Homeless Person Assistance}

\subsection{Case Study I - Studying Noise Complaints by NYC Zip Code}

Invalid values?
Logical inconsistencies? Concerning patterns?
Redundant columns?

Invalid data: closed data earlier than creation date;
latitude/longitude not in NYC; invalid zip code;

Inconsistent data: upper/lower cases; extra spaces; 
open text entries: provide enough major categories and sub categories;

Inefficiently stored data: lat/long versus (lat, long);
intersection street 1 versus cross street 1;
borough could be inferred from zip code;
full name of agency not needed

\section{Improvements} \label{sec:improve}

Data validation: identify invalid data

Data cleaning: convert to all lower/upper cases;

Data organization: put zipcode/borough in a separate table; verify intersection
== cross; remove redundant columns (and compare how much space saved)

\section{Protocol Suggestions} \label{sec:protocol}

\section{Discussion} \label{sec:disc}

\bibliographystyle{asa}
\bibliography{ref}

\end{document}

\end{document}