### 2025-02-07 data_anomalies/code/data_anomaly_presentation.qmd
Establishing a new folder for the 2025 NYC Open Data Week
Initial updload to Git for the work associated with the 2025 NYC Open Data Week presentations.  Expect to delete many of the files in the "function" folder over time.

---
title: "Identifying Data Anomalies"
author: 
  - "David Tussey (formerly w/DoITT)"
  - "Dr. Jun Yan (Dept of Statistics, UConn)"
format: revealjs
---

## Introduction

-   NYC 311 Service Requests
-   October through December 2024
-   Over 900,000 service requests
-   41 fields of information

## What is Data Hygiene?

Key aspects we'll examine:

-   Missing Values
-   Invalid Values
-   Inconsistent Representations
-   Structural Issues

## Missing Values Analysis {background-color="#f8f9fa"}

Is our data missing any values?

Let's analyze key fields:

::: callout-tip
## Interactive Analysis

[Launch Missing Values Check](http://127.0.0.1:4749)
:::

## Interpreting Missing Values

What we found:

-   `incident_zip`: About 1% blank values
-   `descriptor`: Mix of blank and "Missing" values
-   `resolution_description`: N/A values common
-   Several fields with multiple missing value types

## Next Analysis: ZIP Code Validation {background-color="#f8f9fa"}

Are all ZIP codes valid for NYC?

::: callout-tip
## Interactive Analysis

[Launch ZIP Code Check](http://127.0.0.1:6353)
:::

## Future Analyses

Additional checks we can run:

-   Date field consistency
-   Borough validation
-   Geographic coordinate validation
-   Category standardization

## Recommendations

Based on our findings:

1.  Standardize how missing values are recorded
2.  Add input validation for ZIP codes
3.  Review data collection process
4.  Document data quality issues found